### 2011-12-06 paper/abstract.tex
m

\documentclass{sig-alternate}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\title{Boosted Regression Trees for Data-Driven Auto-Tuning}
% Automatic
% of GPU Code
% Just-In-Time Code Specialization

\numberofauthors{3}

\author{
%\alignauthor First Last\titlenote{}\\
\alignauthor First Last\\
\affaddr{Affiliation line 1}\\
\affaddr{Affiliation line 2}\\
\email{anon@mail.com}
% 2nd author
%\alignauthor First Last\titlenote{None}\\
\alignauthor First Last\\
\affaddr{Affiliation line 1}\\
\affaddr{Affiliation line 2}\\
\email{anon@mail.com}
% 3rd author
%\alignauthor First Last\titlenote{None}\\
\alignauthor First Last\\
\affaddr{Affiliation line 1}\\
\affaddr{Affiliation line 2}\\
\email{anon@mail.com}
}

\begin{document}
\maketitle

\begin{abstract}
Auto-tuning is a widely used and effective technique for optimizing a
parametrized GPU code template for a particular type of computation on
particular hardware.
Its drawback is that thorough or exhaustive auto-tuning requires compiling
many kernels, and calling each one many times; this process is slow.
Furthermore, library abstraction boundaries provide operations such as image
filtering and matrix multiplication, which actually correspond to a large set
of potential problem configurations with a wide variety of memory
access patterns and computational bottlenecks.
How can we draw on data from previous auto-tuning of related problems on related hardware to make a
good decision on a novel problem?
This paper presents a machine learning approach to auto-tuning, in
which features of
(a) the current hardware platform,
(b) the kernel configuration
and (c) the problem instance are used to fit a regression model (boosted regression trees) that predicts
how much faster an implementation will be than a reference baseline.
Combinatorial optimization strategies that would normally be applied for auto-tuning can be applied to the regression model to find an implementation much more quickly.
We validate our approach using the filterbank correlation kernel described in [anon].
We find that 0.05 seconds of hill climbing on the regression model can achieve
80\% of the improvement brought by 120 seconds of real auto-tuning.
Our approach is not specific to filterbank correlation, or even to GPU auto-tuning: the approach of using a non-linear regression model on top of simple features applies to a variety of problem types, kernel types, and platforms.
A detailed platform model can provide useful features for
cross-platform auto-tuning, but such a model is not necessary for same-platform autotuning.
\end{abstract}

\end{document}